Active-contours driven registration of diffusion MR images of the brain

### 2015-03-16 Publications/2015-NeuroImage/02-abstract.tex
update figure, updates in manuscript

% -*- root: 00-main.tex -*-
Integrating information from different images is an ubiquitous necessity
  in current neuroimage analyses.
However, oftentimes one of the images is not available or the objective
  function is not adequate or numerically tractable.
We propose a surfaces-to-volume registration method called \emph{regseg}
  that exploits the detailed anatomy extracted from structural images and
  simultaneously segments the target image with subvoxel precision.
We use active contours without edges to search for a deformation field
  of B-Spline functions that optimally maps the surfaces
  onto the data in the target space.
We test \emph{regseg} on four digital phantoms warped with random deformations,
  and conclude that the precision of the method is below the half-voxel spacing.
Finally, we apply \emph{regseg} in the correction for susceptibility-derived
  distortions of 16 real diffusion MRI datasets from the \acrlong*{hcp}
  database.
The results suggest that \emph{regseg} overperformed an extended competing method.
\emph{Regseg} showed a consistent error within the 95\% confidence interval of
  $0.39-0.93mm$ in images of $1.25mm$ isotropic resolution.

>>>>

% -*- root: 00-main.tex -*-
We propose a simultaneous segmentation and registration method called \emph{regseg}
  that uses active contours without edges extracted from structural images.
The contours evolve through a free-deformation field of B-Spline functions that
  optimally maps the surfaces onto the data in the target space.
\emph{Regseg} is intended, but not limited, to correct susceptibility-derived distortions
  in echo-planar imaging.
We proof the concept behind \emph{regseg} on four digital phantoms warped with random
  deformations, and conclude that the precision of the segmentation is subvoxel.
Finally, we demonstrate \emph{regseg} in the correction task on 16 real diffusion MRI
  datasets from the \acrlong*{hcp}.
The results suggest that \emph{regseg} overperformed an extended competing method.
\emph{Regseg} showed a consistent error within the 95\% confidence interval of
  $0.39-0.93mm$ in images of $1.25mm$ isotropic resolution.

### 2015-03-16 Publications/2015-NeuroImage/04-methods.tex
update figure, updates in manuscript

\begin{figure*}
	\includegraphics[width=\linewidth]{figures/figure03}
	\caption{A. The ``cortex'' phantom is a spherical shape with two sulci and an
	  outer crust resembling the cortical folding (left).
	The model is used to generate \gls*{t1} and \gls*{t2} images after warping the
	  contours using a random and plausible transformation $U_{true}^{-1}$ (right).
	B. Visual assessment of the results on the low resolution sets:
	  ``gyrus'' (top-left), ``L-shape'' (top-right), ``ball'' (bottom-left),
	  and ``box'' at (bottom-right).
	In yellow color, the recovered contours after registration are represented.
	Our method showed high accuracy, as demonstrates almost exact location of the registered
    contours with respect to their ground truth position depicted in green.
	Partial volume effect turns segmentation of the sulci a challenging problem with voxel-wise
	  clustering methods, but it is successfully segmented with our method.
	C. Quantitative evaluation of registration error in terms of average Hausdorff distance of
	  surfaces at high (left) and low (right) resolutions, demonstrating that the error is
	  consistently below the voxel size.
	  }\label{fig:phantom}
\end{figure*}

>>>>

\begin{figure*}
    \centering
    \resizebox{\textwidth}{!}{%
      \begin{tikzpicture}
        \node[inner sep=5pt, draw=black!40](fig03a) at (0,14)
          {\includegraphics[width=\linewidth]{figures/figure03-a}};
        \node[inner sep=5pt, draw=black!40](fig03b) at (0,7.6)
          {\includegraphics[width=\linewidth]{figures/figure03-b}};
        \node[inner sep=5pt, draw=black!40](fig03c) at (0,0)
          {\includegraphics[width=\linewidth]{figures/figure03-c}};
        \node[circle, text=black!75] at (-8.8,16.2) {\Huge \textbf{A}};
        \node[circle, text=black!75] at (-8.8,10.05) {\Huge \textbf{B}};
        \node[circle, text=black!75] at (-8.8,3.4) {\Huge \textbf{C}};
      \end{tikzpicture}
    }%
	\caption{A. The ``cortex'' phantom is a spherical shape with two sulci and an
	  outer crust resembling the cortical folding (left).
	The model is used to generate \gls*{t1} and \gls*{t2} images after warping the
	  contours using a random and plausible transformation $U_{true}^{-1}$ (right).
	B. Visual assessment of the results on the low resolution sets:
	  ``gyrus'' (top-left), ``L-shape'' (top-right), ``ball'' (bottom-left),
	  and ``box'' at (bottom-right).
	In yellow color, the recovered contours after registration are represented.
	Our method showed high accuracy, as demonstrates almost exact location of the registered
    contours with respect to their ground truth position depicted in green.
	Partial volume effect turns segmentation of the sulci a challenging problem with voxel-wise
	  clustering methods, but it is successfully segmented with our method.
	C. Quantitative evaluation of registration error in terms of average Hausdorff distance of
	  surfaces at high (left) and low (right) resolutions, demonstrating that the error is
	  consistently below the voxel size.
	  }\label{fig:phantom}
\end{figure*}

### 2015-03-16 Publications/2015-NeuroImage/06-discussion.tex
update figure, updates in manuscript

\paragraph*{Application on real data}
After a preliminary experiment to demonstrate the aptness of our segmentation model
  in section \nameref{sec:res_model_and_metric}, we evaluated \emph{regseg} in
  a real environment, using the experimental framework presented in \autoref{fig:evworkflows}.
We processed 16 subjects from the \gls*{hcp} database with both \emph{regseg}
  and an in-house replication of a competing method.
This alternative approach solves distortion using \emph{b0}-to-\gls*{t2} registration.
Our method presented significantly different results (ANOVA tests are reported in
  \autoref{tab:results_real}), and the 95\% CIs revealed the high accuracy of
  \emph{regseg}.
The aggregated CI for \emph{regseg} was $[0.39 - 0.93] mm$, whereas the competing method
  yielded an aggregated CI of $[2.36 - 3.40] mm$.
Even though we carefully configured the competing method by using the same algorithm and
  extending the same settings from a widely used software, cross-comparison
  experiments are prone to the so-called instrumentation bias \citep{tustison_instrumentation_2013}.
Therefore, these results do not show that the \emph{regseg} software \emph{is} better than the
  competing method, but they are communicated to proof that \emph{regseg} is a solid option to
  correct \gls*{dmri} data for distortions, and any other potential application.

>>>>

\paragraph*{Application on real data}
After a preliminary experiment to demonstrate the aptness of our segmentation model
  in section \autoref{sec:res_model_and_metric}, we evaluated \emph{regseg} in
  a real environment, using the experimental framework presented in \autoref{fig:evworkflows}.
We processed 16 subjects from the \gls*{hcp} database with both \emph{regseg}
  and an in-house replication of a competing method.
This alternative approach solves distortion using \emph{b0}-to-\gls*{t2} registration.
Our method presented significantly different results (ANOVA tests are reported in
  \autoref{tab:results_real}), and the 95\% CIs revealed the high accuracy of
  \emph{regseg}.
The aggregated CI for \emph{regseg} was $[0.39 - 0.93] mm$, whereas the competing method
  yielded an aggregated CI of $[2.36 - 3.40] mm$.
Even though we carefully configured the competing method by using the same algorithm and
  extending the same settings from a widely used software, cross-comparison
  experiments are prone to the so-called instrumentation bias \citep{tustison_instrumentation_2013}.
Therefore, these results do not show that the \emph{regseg} software \emph{is} better than the
  competing method, but they are communicated to proof that \emph{regseg} is a solid option to
  correct \gls*{dmri} data for distortions, and any other potential application.
